\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2012-10-22 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added "Cartesian" to the name of the package and to the manual

\begin{document}
\begin{center}
\thispagestyle{empty}%
\vspace*{-1.8truecm}
\noindent\makebox[\textwidth]{%
\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d-cover.pdf}
}
\end{center}
%
\title{\textbf{SPECFEM3D}\\
\textbf{User Manual}}

@@

\section*{Authors}
The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\

@@

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\

@@

All SPECFEM3D software is written in Fortran90 with full portability
in mind, and conforms strictly to the Fortran95 standard. It uses
no obsolete or obsolescent features of Fortran77. The package uses
parallel programming based upon the Message Passing Interface (MPI)
\citep{GrLuSk94,Pac97}.\\

@@

SPECFEM3D includes coupled fluid-solid domains and
adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
To accommodate visco-plastic or non-linear rheologies, the reader can refer to
the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.

@@

If you use SPECFEM3D for your own research, please cite at least one
of the following articles:

@@

The SPECFEM3D software package comes in a gzipped tar ball. In the
directory in which you want to install the package, type
\begin{lyxcode}
{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
\end{lyxcode}
The directory \texttt{\small SPECFEM3D/} will then contain
the source code.\\

@@

The SPECFEM3D software package relies on the SCOTCH library to partition meshes created with CUBIT.
Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.

@@

The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
\begin{description}
\item [1.] setting up a hexahedral mesh with CUBIT,
\item [2.] exporting the CUBIT mesh into a SPECFEM3D file format and
\item [3.] partitioning the SPECFEM3D mesh files for a chosen number of cores.
\end{description}

Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}. We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two. This example can be compared for validation against the solutions provided in subdirectory ~\\
\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}

@@

The SPECFEM3D software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
the execution of the solver so it is referred to as a static mapping.

@@

In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
During the configuration process, the script tries to find existing SCOTCH installations.
In case your system has no pre-existing SCOTCH installation,
we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
in directory
\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
This version gets bundled with the compilation of the SPECFEM3D package if no libraries could have been found.
If this automatic compilation of the SCOTCH libraries fails,
please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you have correctly specified the path
of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
\\

@@

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/}, and
     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition. These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
\end{itemize}
Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.

@@

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/meshfem3D\_files/}.
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.

@@

These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{in\_out\_files/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.

@@

Mesher output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements.
Alternatively, output can be directed
to the screen instead by uncommenting a line in \texttt{constants.h}:
\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

@@

To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{in\_out\_files/OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh). To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.

@@

\caption{Schematic workflow for a SPECFEM3D simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}

@@

In the main directory, type
\begin{lyxcode}
{\small make xgenerate\_databases}{\small \par}
\end{lyxcode}
Input for the program is provided through
the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{in\_data\_files/}. Please note that
\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.

@@

Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
located in the subdirectory \texttt{in\_data\_files/}:

@@

\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz} in directory \texttt{in\_data\_files/}. See Section \ref{sec:Using-tomographic},

@@

the file system (\texttt{timestamp{*}} files in the \texttt{in\_out\_files/OUTPUT\_FILES}

@@

information about the source time function in the file \texttt{in\_out\_files/OUTPUT\_FILES/plot\_source\_time\_function.txt}.

@@

The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed. In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
\texttt{in\_out\_files/OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions. Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.

@@

The solver needs three input files in the \texttt{in\_data\_files/} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}

@@

\begin{lyxcode}
{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
\end{lyxcode}
The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{in\_data\_files/STATIONS}
to exclude stations that are not located within the region given in
the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
The filtered file is called \texttt{in\_data\_files/STATIONS\_FILTERED}.

Solver output is provided in the \texttt{in\_out\_files/OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{constants.h}:

@@

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}' files in the ~\\
\texttt{\small in\_out\_files/OUTPUT\_FILES} directory.
These tiny files look something like this:

@@

\begin{lyxcode}
utils/change\_simulation\_type.pl~-f~
\end{lyxcode}
and then collect the recorded seismograms at all the stations given
in \texttt{in\_data\_files/STATIONS}.

\item Then select the stations for which you want to compute the time-reversed
adjoint sources and run the adjoint simulation, and compile them into
the \texttt{in\_data\_files/STATIONS\_ADJOINT} file, which has the same format
as the regular \texttt{in\_data\_files/STATIONS} file.

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in the \texttt{in\_out\_files/SEM/} directory
with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
are the station name and network code given in the \texttt{in\_data\_files/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending on
your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
In the \texttt{in\_data\_files/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

@@

\begin{itemize}
\item These adjoint seismograms are recorded at the locations of the original
earthquake sources given by the \texttt{in\_data\_files/CMTSOLUTION} file, and
have names of the form \texttt{S?????.NT.S??.sem} for the six-component
strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
and ~\\
\texttt{S?????.NT.BX?.sem} for the three-component displacements
(\texttt{BXN,BXE,BXZ}) recorded at these locations.
\item \texttt{S?????} denotes the source number; for example, if the original
\texttt{CMTSOLUTION} provides only a point source, then the seismograms
collected will start with \texttt{S00001}.
\item These adjoint seismograms provide critical information for the computation
of the gradient of the misfit function.
\end{itemize}
\end{enumerate}

@@

With the successful forward simulation and the adjoint source ready
in the \texttt{in\_out\_files/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
and rerun the solver.

@@

Besides earthquake simulations, SPECFEM3D includes functionality for seismic noise tomography as well.
In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
in order to understand noise simulations from a theoretical perspective.

@@

As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
Unless otherwise specified, those input files should be located in directory \texttt{in\_data\_files/}.\\

@@

\texttt{*************************************************************}\\
\texttt{the source time function has been saved in:}\\
\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
\texttt{S\_squared should be put into directory:}\\
\texttt{in\_out\_files/NOISE\_TOMOGRAPHY/ in the SPECFEM3D package}\\

@@

\item
Create a new directory in the SPECFEM3D package, name it as \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.
We will add some parameter files later in this folder.

\item
Put the Matlab-generated-file \texttt{S\_squared} in \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/}.

That's to say, you will have a file \texttt{in\_out\_files/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D package.

>>>>

\begin{document}
\begin{center}
\thispagestyle{empty}%
\vspace*{-1.8truecm}
\noindent\makebox[\textwidth]{%
\noindent\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover.pdf}
}
\end{center}
%
\title{\textbf{SPECFEM3D Cartesian}\\
\textbf{User Manual}}

@@

\section*{Authors}
The SPECFEM3D Cartesian package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998, based on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe (IPGP) in Paris, France from 1995 to 1997.\\

@@

The software package SPECFEM3D Cartesian simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\

@@

All SPECFEM3D Cartesian software is written in Fortran90 with full portability
in mind, and conforms strictly to the Fortran95 standard. It uses
no obsolete or obsolescent features of Fortran77. The package uses
parallel programming based upon the Message Passing Interface (MPI)
\citep{GrLuSk94,Pac97}.\\

@@

SPECFEM3D Cartesian includes coupled fluid-solid domains and
adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
To accommodate visco-plastic or non-linear rheologies, the reader can refer to
the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.

@@

If you use SPECFEM3D Cartesian for your own research, please cite at least one
of the following articles:

@@

The SPECFEM3D Cartesian software package comes in a gzipped tar ball. In the
directory in which you want to install the package, type
\begin{lyxcode}
{\small tar~-zxvf~SPECFEM3D\_V2.1.0.tar.gz}{\small \par}
\end{lyxcode}
The directory \texttt{\small SPECFEM3D/} will then contain
the source code.\\

@@

The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.

@@

The basic steps in creating a load-balanced, partitioned mesh with CUBIT are:
\begin{description}
\item [1.] setting up a hexahedral mesh with CUBIT,
\item [2.] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format and
\item [3.] partitioning the SPECFEM3D Cartesian mesh files for a chosen number of cores.
\end{description}

Examples are provided in the SPECFEM3D Cartesian package in the subdirectory \texttt{examples/}. We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

For the installation and handling of the CUBIT meshing tool suite, please refer to the CUBIT user manual and documentation.
In order to give you a basic understanding of how to use CUBIT for our purposes, examples are provided in the SPECFEM3D Cartesian package in the subdirectory \texttt{examples/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}]  Combines two different, elastic material volumes and creates a refinement layer between the two. This example can be compared for validation against the solutions provided in subdirectory ~\\
\texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose material properties will have to be read in from a tomographic model file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}

@@

The SPECFEM3D Cartesian software package performs large scale simulations in a parallel 'Single Process Multiple Data' way.
The spectral-element mesh created with CUBIT needs to be distributed on the processors. This partitioning is executed once and for all prior to
the execution of the solver so it is referred to as a static mapping.

@@

In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
During the configuration process, the script tries to find existing SCOTCH installations.
In case your system has no pre-existing SCOTCH installation,
we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
in directory
\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
This version gets bundled with the compilation of the SPECFEM3D Cartesian package if no libraries could have been found.
If this automatic compilation of the SCOTCH libraries fails,
please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you have correctly specified the path
of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
\\

@@

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/}, and
     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition. These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
\end{itemize}
Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.

@@

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding of how
the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.

@@

These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.

@@

Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Please note that the mesher suggests a time step \texttt{DT} to run
the solver with. The mesher output file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements.
Alternatively, output can be directed
to the screen instead by uncommenting a line in \texttt{constants.h}:
\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

@@

To control the quality of the mesh, check the standard output (either on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt}) and analyze the skewness histogram of mesh elements to make sure no element has a skewness above approximately 0.75, otherwise the mesh is of poor quality (and if even a single element has a skewness value above 0.80, then you must definitely improve the mesh). To draw the skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.

@@

\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable \texttt{xgenerate\_databases} creates the GLL mesh points and assigns specific model parameters.}

@@

In the main directory, type
\begin{lyxcode}
{\small make xgenerate\_databases}{\small \par}
\end{lyxcode}
Input for the program is provided through
the main parameter file \texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}. Please note that
\texttt{xgenerate\_databases} must be called directly from the \texttt{bin/} directory, as most of the binaries of the package.

@@

Before running \texttt{xgenerate\_databases}, a number of parameters need to be set in the main parameter \texttt{Par\_file}
located in the subdirectory \texttt{DATA}:

@@

\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses a tomographic model file \texttt{tomographic\_model.xyz}
in directory \texttt{DATA}. See Section \ref{sec:Using-tomographic},

@@

the file system (\texttt{timestamp{*}} files in the \texttt{OUTPUT\_FILES}

@@

information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.

@@

The database generator \texttt{xgenerate\_databases}, as well as the internal mesher \texttt{xmeshfem3D}, are trying
to evaluate the value of $\Delta t$ for empirically chosen Courant numbers $C \sim 0.3$.
If you used the mesher \texttt{xmeshfem3D} to generate your mesh, you should set the value suggested in \texttt{OUTPUT\_FILES/output\_mesher.txt} file, which is created after the mesher completed. In case you used CUBIT to create the mesh, you might use an arbitrary value when running \texttt{xgenerate\_databases} and then use the value suggested in the ~\\
\texttt{OUTPUT\_FILES/output\_mesher.txt} file after the database generation completed.
Note that the implemented Newmark time scheme uses this time step globally, thus your simulations become more expensive for very small mesh elements in high wave-speed regions. Please be aware of this restriction when constructing your mesh in Chapter~\ref{cha:Mesh-Generation}.

@@

The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}

@@

\begin{lyxcode}
{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
\end{lyxcode}
The solver \texttt{xspecfem3D} filters the list of stations in file \texttt{DATA/STATIONS}
to exclude stations that are not located within the region given in
the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN} and \texttt{LATITUDE\_MAX}
and between \texttt{LONGITUDE\_MIN} and \texttt{LONGITUDE\_MAX}).
The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.

Solver output is provided in the \texttt{OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{constants.h}:

@@

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}' files in the ~\\
\texttt{\small OUTPUT\_FILES} directory.
These tiny files look something like this:

@@

\begin{lyxcode}
utils/change\_simulation\_type.pl~-f~
\end{lyxcode}
and then collect the recorded seismograms at all the stations given
in \texttt{DATA/STATIONS}.

\item Then select the stations for which you want to compute the time-reversed
adjoint sources and run the adjoint simulation, and compile them into
the \texttt{DATA/STATIONS\_ADJOINT} file, which has the same format
as the regular \texttt{DATA/STATIONS} file.

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in the \texttt{OUTPUT\_FILES/SEM/} directory
with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA}, \texttt{NT}
are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending on
your sampling rate (see Appendix~\ref{cha:channel-codes} for further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

@@

\begin{itemize}
\item These adjoint seismograms are recorded at the locations of the original
earthquake sources given by the \texttt{DATA/CMTSOLUTION} file, and
have names of the form \texttt{S?????.NT.S??.sem} for the six-component
strain tensor (\texttt{SNN,SEE,SZZ,SNE,SNZ,SEZ}) at these locations,
and ~\\
\texttt{S?????.NT.BX?.sem} for the three-component displacements
(\texttt{BXN,BXE,BXZ}) recorded at these locations.
\item \texttt{S?????} denotes the source number; for example, if the original
\texttt{CMTSOLUTION} provides only a point source, then the seismograms
collected will start with \texttt{S00001}.
\item These adjoint seismograms provide critical information for the computation
of the gradient of the misfit function.
\end{itemize}
\end{enumerate}

@@

With the successful forward simulation and the adjoint source ready
in the \texttt{OUTPUT\_FILES/SEM/} directory, set \texttt{SIMULATION\_TYPE = 3} and \texttt{SAVE\_FORWARD
= .false.} in the \texttt{Par\_file} (you can use \texttt{change\_simulation\_type.pl -b}),
and rerun the solver.

@@

Besides earthquake simulations, SPECFEM3D Cartesian includes functionality for seismic noise tomography as well.
In order to proceed successfully in this chapter, it is critical that you have already familiarized yourself
with procedures for meshing (Chapter \ref{cha:Mesh-Generation}),
creating distributed databases (Chapter \ref{cha:Creating-Distributed-Databases}),
running earthquake simulations (Chapters \ref{cha:Running-the-Solver})
and adjoint simulations (Chapter \ref{cha:Adjoint-Simulations}).
Also, make sure you read the article `Noise cross-correlation sensitivity kernels' \citep{trompetal2010},
in order to understand noise simulations from a theoretical perspective.

@@

As usual, the three main input files are crucial: \texttt{Par\_file}, \texttt{CMTSOLUTION} and \texttt{STATIONS}.
Unless otherwise specified, those input files should be located in directory \texttt{DATA/}.\\

@@

\texttt{*************************************************************}\\
\texttt{the source time function has been saved in:}\\
\texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be different)\\
\texttt{S\_squared should be put into directory:}\\
\texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian package}\\

@@

\item
Create a new directory in the SPECFEM3D Cartesian package, name it as \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.
We will add some parameter files later in this folder.

\item
Put the Matlab-generated-file \texttt{S\_squared} in \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/}.

That's to say, you will have a file \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/S\_squared} in the SPECFEM3D Cartesian package.

\item
Create a file called \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/irec\_master\_noise}. Note that this file is located
in directory \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/} as well. In general, all noise simulation related parameter
files go into that directory.
\texttt{irec\_master\_noise} contains only one integer, which is the ID
of the `master' receiver. For example, if this file contains 5, it means that the fifth receiver
listed in \texttt{DATA/STATIONS} becomes the `master'. That's why we mentioned previously that
the order of receivers in \texttt{DATA/STATIONS} is important.
